---
title: "UPDATED Supplementary9_Poepoe_et_al_RoyalB_Figures_and_Analysis"
output:
  html_document: default
  pdf_document: default
editor_options:
  markdown:
    wrap: 72
---

**Poepoe et al. Supplementary 9**
**Quantitative Data Analysis**

Thank you to the following statisticians for their review of the Linear Mixed Models section: Etaash Katiyar (Stanford University), Wesley Bian (Stanford University), Louis Reeve Stanley Davis (Stanford University), Jules Wyman (Washington University)

**Set Up**

```{r setup, include = FALSE}

#Clear workspace
rm(list=ls()) 

# Set a default CRAN mirror
options(repos = c(CRAN = "https://cloud.r-project.org"))

```

```{r}

#Set working directory
setwd("~/Desktop") 
getwd()

#For plotting
install.packages('ggtext')
install.packages('showtext')

#Load libraries
library(ggplot2)
library(dplyr)
#library(ggtext)     # Text rendering
library(showtext)   # Custom fonts
library(sysfonts)

```

**Analysis for Size at Spawn (All Islands)**

## PART 1: Statistical testing against a theoretical distribution

**Input:**

1.  *10a_raw-data.csv*

**Output:**

1.  *10a_all-analysis.csv*

2.  *10a_p\<0.001.csv*

Load in CSV

```{r}
setwd("/Users/marina-work/Desktop")
fig3data <-read.csv("10a_raw-data.csv") #Raw data used for making Figure 3
head(fig3data)

#Rename Columns
#fig3data <- fig3data %>% rename(
#  Date = Date..mm.dd.yyyy.,
#  obs_FL = Observed.Fork.Length..mm.,
#)

#Format Columns
fig3data$Trip <- paste(fig3data$Date, fig3data$Island, sep = "_")
fig3data$Date <- as.Date(fig3data$Date,"%m/%d/%Y")
fig3data$obs_FL <- as.numeric(fig3data$obs_FL)
fig3data$Sex <- as.factor(fig3data$Sex)
fig3data$Island <- as.factor(fig3data$Island)

#Format data
fig3data$Year <- format(fig3data$Date, '%Y')
fig3data$Month <- format(fig3data$Date, '%m')

#Separate Males and Females
F_only <- fig3data %>%  filter(Sex == "Female")
M_only <- fig3data %>%  filter(Sex == "Male")


```

Summarize data

```{r}

fig3table <- fig3data %>% 
  group_by(Sex, Trip) %>% #create averages per trip
  reframe(
    Date = Date,
    Year = Year,
    Island = Island,
    Sex = Sex,
    obs_FL = obs_FL,
    n= n(),
) %>% 
  group_by(Sex, Trip) %>% 
  mutate(
  k = mean(obs_FL)
  ) %>% 
  ungroup() %>%
  select(Trip, Date, Year, Island, Sex, n, k) %>% 
  distinct()

```

**Introducing three modeled scenarios**

*Least Conservative (LC): 68% of data falls +/- 1 standard deviation
from the mean*

*Middle Conservative (LC): 95% of data falls +/- 2 standard deviations
from the mean*

*Highly Conservative (LC): 99.7% of data falls +/- 3 standard deviations
from the mean*

Adding expected values and setting scenario parameters

```{r}
#Expected values
##Males size range (Rm): 200-250mm (mean = 225mmm)
##Females size range (Rf) = 300-400mm (mean = 350mm)
fig3table$Expected_mean <- 0
fig3table$Expected_mean[fig3table$Sex=='Female'] <- 350
fig3table$Expected_mean[fig3table$Sex=='Male'] <- 225


## Calculating sigma for three scenarios where the range of male/female sizes is one, two and three standard deviations from the mean. 

### Example: For LC (females), the range 400-300mm = 100mm. 100mm is +/- one standard deviation = two standard deviations total. Therefore sigma = 100mm/2 = 50mm. 

### Example: For LC (males), the range 250-200mm = 50mm is +/- one standard deviation, so two standard deviations total. Therefore sigma = 50mm/2 = 25mm. 

## Three scenarios:
### Least Conservative (LC): Rm and Rf are one standard deviation (sigma = 50 for females and 25 for males)	
### Middle Conservative (MC): Rm and Rf are two standard deviations (sigma = 25 for females and 12.5 for males)
### Highly Conservative (HC): Rm and Rf are two standard deviations (sigma = 16.7 for females and 8.3 for males)

#Create empty columns
fig3table$LC_sigma <- 0
fig3table$MC_sigma <- 0
fig3table$HC_sigma <- 0

#Assign parameter values
fig3table$LC_sigma[fig3table$Sex=='Female'] <- 50
fig3table$LC_sigma[fig3table$Sex=='Male'] <- 25
fig3table$MC_sigma[fig3table$Sex=='Female'] <- 25
fig3table$MC_sigma[fig3table$Sex=='Male'] <- 12.5
fig3table$HC_sigma[fig3table$Sex=='Female'] <- 16.7
fig3table$HC_sigma[fig3table$Sex=='Male'] <- 8.3

#Make sure column is formatted correctly
fig3table$n <- as.numeric(fig3table$n)

```

**Calculate Absolute Z-score, Probability Density Function and 2-tailed
P-value**

```{r}
#Absolute Z score
## absolute value (observed mean (k) - expected mean)/(sigma / sqrt(n))

fig3analysis <- fig3table %>% mutate (
  sqrt_n = sqrt(n),
  means_difference = k-Expected_mean,
  LC_absZ = abs(means_difference/(LC_sigma/(sqrt_n))),
  MC_absZ = abs(means_difference/(MC_sigma/(sqrt_n))),
  HC_absZ = abs(means_difference/(HC_sigma/(sqrt_n)))
  )


#Probability Density Function
## apply the probability density function to the absolute z-score

fig3analysis <- fig3analysis %>% mutate (
  LC_pdf = pnorm(LC_absZ,mean = 0, sd = 1, lower.tail=TRUE),
  MC_pdf = pnorm(MC_absZ,mean = 0, sd = 1, lower.tail=TRUE),
  HC_pdf = pnorm(MC_absZ,mean = 0, sd = 1, lower.tail=TRUE)
  )

#Two-tailed p-value
## 2*(1- Probability density function)

fig3analysis <- fig3analysis %>% mutate (
  LC_p_value = 2*(1-LC_pdf),
  MC_p_value = 2*(1-MC_pdf),
  HC_p_value = 2*(1-HC_pdf)
  )
  
#Save csv
#write.csv(fig3analysis, "10a_all-analysis.csv")
  

```

**Looking at which fish observations are larger and smaller than
expected under the three scenarios**

```{r}
#Least conservative scenario
LC <- fig3analysis %>% filter(
    LC_p_value <= 0.01
  ) %>%  group_by(Island, Sex, Trip) %>% reframe(
  n_island = n,
  size_difference = mean(k) - Expected_mean,
  mean_FL=mean(k),
  Year = Year,
  Date = Date
) %>% distinct()

#Middle conservative scenario
MC <- fig3analysis %>% filter(
    MC_p_value <= 0.01
  ) %>%  group_by(Island, Sex, Trip) %>% reframe(
  n_island = n,
  size_difference = mean(k) - Expected_mean,
  mean_FL=mean(k),
  Year = Year,
  Date = Date
) %>% distinct()

#Highly conservative scenario
HC <- fig3analysis %>% filter(
    HC_p_value <= 0.01
  ) %>%  group_by(Island, Sex, Trip) %>% reframe(
  n_island = n,
  size_difference = mean(k) - Expected_mean,
  mean_FL=mean(k),
  Year = Year,
  Date = Date
) %>% distinct()

#Interpretation:
#Maui females are unexpectedly small
#Oahu males are unexpectedly large

#Kauai males are unexpectedly large
#Hawai'i males are unexpectedly large
#Molokai males are unexpectedly small 


#save csv
LC$scenario <- "LC"
MC$scenario <- "MC"
HC$scenario <- "HC"

Fig3analysis_key <- rbind(LC, MC, HC)
  
write.csv(Fig3analysis_key, "10a_p<0.001.csv")


```

## PART 2: Linear models
Results from PART 1: 

1. In February and April 2024 on Maui fish are female at smaller sizes (270-300mm, instead of over 300mm). 

2. In September 2024 on 'Oahu fish are male at larger sizes (350-380mm) compared to generational knowledge (Figure 3) and previous records

Question: Are the findings from PART 1 confirmed through a modeling approach? Is moi size-at sex different across islands and years? What are the trends?

Approach: Look at what factors contribute to differences in moi fork length using generalized linear mixed models. 

Variables: Sex, Island, Month, Day-of-year, Year, Day count


**Input:**

1.  *10a_raw-data.csv*

**Output:**

```{r}

#Load libraries
library(lme4)
library(emmeans)
library(glmmTMB)
library(DHARMa)
library(MuMIn)
library(ggplot2)
library(dplyr)
library(splines)
library(AICcmodavg)
install.packages("diptest"); library(diptest)

setwd("/Users/marina-work/Desktop")
getwd()
lmm_data <- read.csv("10a_raw-data.csv")

#Format data
str(lmm_data)
lmm_data$Sex <- as.factor(lmm_data$Sex)
lmm_data$Island <- as.factor(lmm_data$Island)
lmm_data$Year <- as.factor(lmm_data$Year)
#lmm_data$Month <- as.factor(lmm_data$Month)
lmm_data$Month_f <- as.factor(lmm_data$Month)
lmm_data$Date <- as.Date(lmm_data$Date, "%Y-%m-%d")
lmm_data$Day_of_year <- format(lmm_data$Date, "%j")
lmm_data$Day_of_year <- as.numeric(lmm_data$Day_of_year)

origin_date <- as.Date("2019-01-01")

lmm_data <- lmm_data %>%
  mutate(
    Day_count = as.numeric(Date - origin_date) + 1
  ) 

```


#Basic descriptions/spread of data
```{r}
with(lmm_data, table(Sex, Island)) #All sexes were sampled on all islands
with(lmm_data, table(Sex, Month)) #not all sexes were sampled for all Months
with(lmm_data, table(Island, Month)) #not all Islands were sampled for all Months

table(lmm_data$Year, lmm_data$Island)
table(lmm_data$Year, lmm_data$Sex)


```



#Distribution of outcome variable (Fork Length)
```{r}
#check for multi-modal distribution which is what we would expect for the different sexes
hist(lmm_data$obs_FL) 
ggplot(lmm_data, aes(x = obs_FL)) +
  geom_density(fill = "skyblue", alpha = 0.7, color = "black") +
  labs(
    title = "Density Plot of Observed Fork Length (mm)",
    subtitle = "Visual check for distinct peaks",
    x = "Observed Fork Length",
    y = "Density"
  ) +
  theme_minimal(base_size = 14)

#dip_test_result <- dip.test(obs_fl_complete) #yes, fork length is multi-modally distributed. 

```

#Which variables to include
Single variable: Sex, Island, Year, Month, Date
Interactions: Sex:Island, Sex:Month, Island:Month
Note: Cannot include 3-way interaction because of gaps in the data


#Which distribution is most appropriate: Gamma vs. Normal
```{r}
#CRITERIA:
#Are the residuals normally distributed?
#Which distribution results in a better fitted model?
#Note: Not enough data to fit interactive effect of Island:Month

#Gamma distribution
m1_gamma <- glmmTMB(obs_FL ~ Sex + Island + Year + Month + Sex:Island + Sex:Month + Island:Month, data = lmm_data, family=Gamma (link ="log") ) 
plot(simulateResiduals(m1_gamma))
summary(m1_gamma)
AICc(m1_gamma) #1178

#Normal distribution (gaussian)
m1_normal <- glmmTMB(obs_FL ~ Sex + Island + Year + Month + Sex:Island + Sex:Month + Island:Month, data = lmm_data, family=gaussian ) 
summary(m1_normal)
plot(simulateResiduals(m1_normal))
AICc(m1_normal) #1174

#CONCLUSION: Use normal distribution
#Models perform about the same, but we would expect these data to be normally distributed (normal distribution is more natural when talking about weights, lengths, heights), versus the gamma distribution was created for modeling wait times/rates of arrival. 


```

#Modeling time of year in months (categorical) vs. continuously
```{r}

#Month as an integer
m1_month <- glmmTMB(obs_FL ~ Sex + Island + Year + Month + Sex:Island + Sex:Month + Island:Month, data = lmm_data, family=gaussian) 
plot(simulateResiduals(fittedModel = m1_month))
summary(m1_month)
AICc(m1_month) #1174

#Time of year continuous: spline over day of year (df=12) 
m1_day_continuous <- glmmTMB(obs_FL ~ Sex + Island + Year + ns(Day_of_year, df=12 ) + Sex:Island + Sex:Month + Island:Month, data = lmm_data, family=gaussian) 
plot(simulateResiduals(m1_day_continuous))
summary(m1_day_continuous) #this confirms that there is a 12x per year cycle that helps to explain variation in fork length. This cycle does not exactly match to month (otherwise both models would perform equally well and this one performs better). 
AICc(m1_day_continuous) #1140

#Overall all models confirm that the sampled moi fork lengths vary across the time of year

#We decided to still use "Month" as a variable in our models to aid interpretability and because the hypotheses we are testing are specified by month. e.g "Male moi are larger in September on 'Oahu"

```


#Testing whether a mixed model is more appropriate (random effects): No
```{r}
#We are interested in the individual effects of Sex, Island, Month and Year so it does not make sense to model them as random variables. 

#comparing linear model vs linear mixed model
m1_normal <- glmmTMB(obs_FL ~ Sex + Island + Year + Month + Sex:Island + Sex:Month + Island:Month, data = lmm_data, family=gaussian) 
plot(simulateResiduals(m1_normal))
summary(m1_normal) 
AICc(m1_normal) #1174


#linear mixed model
m1_yr_random <- glmmTMB(obs_FL ~ Sex + Island + (1|Year) + Month + Sex:Island + Sex:Month + Island:Month, data = lmm_data, family=gaussian ) 
plot(simulateResiduals(m1_yr_random)) 
summary(m1_yr_random)
AICc(m1_yr_random) #1170 (if Island:Month is removed)


#Modeling year as a fixed variable because we are interested in looking at the variation between years

```

#Model Selection
Evaluate according to AICC (best for small sample sizes)
```{r}

#No interaction effects
m1 <- glmmTMB(obs_FL ~ Sex + Island + Month + Year, data = lmm_data, family = gaussian) 
summary(m1)
plot(simulateResiduals(m1)) 
testDispersion(m1)
AICc(m1)
#aicc:1188
sst <- sum((lmm_data$obs_FL - mean(lmm_data$obs_FL))^2)
sse <- sum((residuals(m1))^2)
r2 <- 1 - sse / sst
#r2: 0.6533425

#Only interaction term of interest
m2 <- glmmTMB(obs_FL ~ Sex + Island + Month + Year + Sex:Island, data = lmm_data, family = gaussian) 
summary(m2)
plot(simulateResiduals(m2))
testDispersion(m2)
AICc(m2)
#aicc: 1173
sst <- sum((lmm_data$obs_FL - mean(lmm_data$obs_FL))^2)
sse <- sum((residuals(m2))^2)
r2 <- 1 - sse / sst
#r2: 0.7222868

#All interaction terms
m3 <- glmmTMB(obs_FL ~ Sex + Island + Month + Year + Sex:Island + Sex:Month + Island:Month , data = lmm_data, family = gaussian) 
summary(m3)
plot(simulateResiduals(m3)) 
testDispersion(m3)
AICc(m3)
#aicc:1174.756
sst <- sum((lmm_data$obs_FL - mean(lmm_data$obs_FL))^2)
sse <- sum((residuals(m3))^2)
r2 <- 1 - sse / sst
r2 #r2: 0.7504138

#All interaction effects have significant terms, and this model (M3) performs well so we will select it over the others.  

```

#Final Model 
```{r}

m_final <- m3

summary(m_final)
plot(simulateResiduals(m_final)) 
testDispersion(m_final)
AICc(m_final) #1174.756

#Calculating an R2 (equivalent)
#Source: https://stats.stackexchange.com/questions/477598/equivalent-of-r-squared-in-generalized-linear-model-regression-results
sst <- sum((lmm_data$obs_FL - mean(lmm_data$obs_FL))^2)
sse <- sum((residuals(m_final))^2)
r2 <- 1 - sse / sst
r2 #r2: 0.7504138


```
#Final Model Interpretation

#1. In February and April 2024 on Maui females were reaching maturity at smaller sizes. 
#2. In September 2024 on 'Oahu males were developing at larger sizes.

```{r}
drop1(m_final, test = "Chisq") 

#Significant (p<0.0001) differences in fork length between sexes among Islands

#Approaching significant differences in the linear relationship between in fork length and Month among Islands (p<0.1)

#Non-significant differences the linear relationship between fork length and Year 

emmeans(m_final, ~ Sex:Island)

##FEMALES
#Create dataframe with emmeans output
emmeans_data <- tibble::tribble(
  ~Sex,    ~Island,      ~emmean, ~SE,  ~df, ~lower.CL, ~upper.CL,
  "Female", "Maui",       303,     8.98, 98,  285,       321,
  "Female", "Kauaʻi",     324,     13.00, 98,  299,       350,
  "Female", "Hawaiʻi",    333,     9.56, 98,  314,       352,
  "Female", "Oʻahu",      371,     17.60, 98,  336,       406,
  "Female", "Molokai",    476,     50.50, 98,  375,       576
)

#Reorder the islands for a cleaner plot (by mean)
emmeans_data$Island <- reorder(emmeans_data$Island, emmeans_data$emmean)

#Plot
females <- ggplot(emmeans_data, aes(x = Island, y = emmean)) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.2,   linewidth = 0.6) + 
  geom_point(size = 4,color = "darkblue") + 
  labs(title = "Female Fork Length - Estimated Mean Value by Island",
       subtitle = "Showing Upper and Lower Confidence Levels",
       x = "Island",
       y = "Estimated Mean (emmean)") +
  theme_minimal(base_size = 12) +
  coord_flip()
#Captured Maui females are smaller than the other islands
 

##MALES
#Create dataframe with emmeans output
emmeans_data_male <- tibble::tribble(
  ~Sex,    ~Island,      ~emmean, ~SE,   ~df, ~lower.CL, ~upper.CL,
  "Male",  "Molokai",    185,     19.00, 98,  148,       223,
  "Male",  "Hawaiʻi",    241,     10.10, 98,  221,       261,
  "Male",  "Kauaʻi",     287,     14.20, 98,  259,       315,
  "Male",  "Maui",       267,     30.10, 98,  207,       326,
  "Male",  "Oʻahu",      344,     13.60, 98,  317,       371
)

#Reorder the islands for a cleaner plot (by mean)
emmeans_data_male$Island <- reorder(emmeans_data_male$Island, emmeans_data_male$emmean)

#Plot
males <- ggplot(emmeans_data_male, aes(x = Island, y = emmean)) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.2,
                linewidth = 0.8) +
  geom_point(size = 4, color = "orange") + 
  labs(title = "Estimated Mean Value by Island (Males)",
       subtitle = "Showing Upper and Lower Confidence Levels",
       x = "Island",
       y = "Estimated Mean (emmean)") +
  theme_minimal(base_size = 14) +
  coord_flip()
#Captured 'Oahu males are larger than the other islands

### Combined plot
combined_data <- rbind(emmeans_data, emmeans_data_male)
#Reorder the islands (by female eemean)
combined_data$Island <- factor(combined_data$Island, levels = reorder(emmeans_data$Island, emmeans_data$emmean) |> levels())

combined_plot <- ggplot(combined_data, aes(x = Island, y = emmean, color = Sex)) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.2, linewidth = 0.6, position = position_dodge(width = 0.5)) +
  geom_point(size = 4, position = position_dodge(width = 0.5)) +
  labs(title = "Moi: Estimated Mean Fork Length by Island and Sex",
       subtitle = "Showing Upper and Lower Confidence Levels",
       x = "Island",
       y = "Moi Mean Fork Length (mm) Linear Model Estimate") +
  theme_minimal(base_size = 10) +
  scale_color_manual(values = c("Female" = "darkblue", "Male" = "orange")) +
  coord_flip()

combined_plot

```

#Additional analysis: targeted linear models for the two main hypotheses, with temporal variables as random effects and month as a factor
1. Are female moi on Maui smaller than females on other islands?
2. Are male moi on 'Oahu larger than males on other islands?

```{r}

females <- lmm_data %>% filter(Sex == "Female")
males <- lmm_data %>% filter(Sex == "Male")

#1. Are female moi on Maui smaller than females on other islands?
m <- glmmTMB(obs_FL ~ Island + (1|Month_f) + (1|Year), data = females, family= gaussian)
summary(m)
#at p<0.0001 significance level the sampled female moi on Maui are smaller than those on the other islands, controlling for the effect of year and month. 

#2. Are male moi on O'ahu larger than males on other islands?
o <- glmmTMB(obs_FL ~ Island + (1|Month_f) + (1|Year), data = males, family = gaussian)
summary(o)
#at p<0.0001 significance level the sampled male moi on O'ahu are larger than those on the other islands, controlling for the effect of year and month. 

```




# Analysis for Spawning Period (Molokai Only)

**Input:**

1.  *10b_raw-data.csv*

**Output:**

1.  *10b_all-analysis.csv*
2.  *10b_Fig4-plot.png*

Load and format data

```{r}

#read in raw data
fig4data <- read.csv("10b_raw-data.csv")

#Rename Columns
#fig4data <- fig4data %>% rename(
#  Day_of_year = Day.of.the.Year,
#  Spawn_evidence = Evidence.of.Spawning,
#  Island = Location,
#)

#Format Columns
fig4data$Date <- as.Date(fig4data$Date,"%m/%d/%Y")
fig4data$Day_of_year <- as.numeric(fig4data$Day_of_year)
fig4data$Island <- as.factor(fig4data$Island)

```

H0: Null Hypothesis Current State of Hawai’i regulation assumes that the
vast majority of spawning activity occurs within the protected
June-August window. Spawning activity outside this window is negligible
or non-existent.

\
Mean (𝛍) 212 day of year Range 92 days

3 Scenarios:

*Range is one Standard Deviation (least conservative) (σ=23.5)*

*Range is two Standard Deviations (middle conservative) (σ=28)*

*Range is three Standard Deviations (highly conservative) (σ=35.9)*

```{r}
#Summarize empirical data
fig4data_analysis <- fig4data %>% 
  group_by(Date, Spawn_evidence) %>% 
  mutate(
  sample_n = n(),
  confirmed_spawn_k = sum(Spawn_evidence == "Yes")
) %>% 
  distinct() %>% 
  group_by(Date) %>% 
  mutate(
    sample_per_day = sum(sample_n)
  )
  

#set parameters

H0_mean_day_of_year <- 212

LC_sd <- 23.5
MC_sd <- 28
HC_sd <- 35.9

```

Analysis

```{r}
###Step 1: Calculation of the probability (p) using normal distribution		
# Question: What the probability to observe spawning at Day X and beyond, under H0?			

#Z score calculations

fig4spawn <- fig4data_analysis %>% filter(Spawn_evidence=="Yes") %>%   
  mutate(
    LC_Z_score = ((Day_of_year)-(H0_mean_day_of_year))/LC_sd,
    MC_Z_score = ((Day_of_year)-(H0_mean_day_of_year))/MC_sd,
    HC_Z_score = ((Day_of_year)-(H0_mean_day_of_year))/HC_sd
  )

#One-tailed probability
#P(Z>zcal) - 1 tail			
fig4spawn <- fig4spawn %>% 
  mutate(
    LC_p = 1-pnorm(abs(LC_Z_score)),
    MC_p = 1-pnorm(abs(MC_Z_score)),
    HC_p = 1-pnorm(abs(HC_Z_score)),
  )

###Step 2: Calculation of p-value 
#Answering the question, what is the probability of observing at least your sample (e.g. 5/6 spawning) under H0 			
#Equation: P(X>=5) = P(X=5)+ P(x=6) with  P(X=k) = C(n, k) * p^k * (1-p)^(n-k)	
#Use the dbinom package in R

fig4spawn <- fig4spawn %>% 
  mutate(
    pvalue_LC = dbinom(x= confirmed_spawn_k, size = sample_n, prob= LC_p),
    pvalue_MC = dbinom(x= confirmed_spawn_k, size = sample_n, prob= MC_p),
    pvalue_HC = dbinom(x= confirmed_spawn_k, size = sample_n, prob= HC_p),
  )

#Save file
write.csv(fig4spawn, "10b_all-analysis.csv")

```

Visualization

```{r}

# Define the x-axis range for the density curves
x_axis_days_of_year <- 1:365

# Find the absolute maximum density across ALL 3 curves
Molokai_max_density <- max(
  dnorm(x_axis_days_of_year, mean = H0_mean_day_of_year, sd = LC_sd),
  dnorm(x_axis_days_of_year, mean = H0_mean_day_of_year, sd = MC_sd),
  dnorm(x_axis_days_of_year, mean = H0_mean_day_of_year, sd = HC_sd)
)

# Maximum frequency across ALL data
Molokai_max_freq <- max(fig4spawn$sample_n, na.rm = TRUE)

# Scaling factor to align the two y-axes
height_multiplier <- 0.9
scaling_factor <- (Molokai_max_freq / Molokai_max_density) / height_multiplier

# --- 3. Plotting with ggplot2 ---

# Define colors for the plot elements
color_curve_line <- "#0077b6"
color_curve_fill <- "aliceblue"
color_spawn_yes <- "darkblue"
color_spawn_no <- "grey"


# Create the plot
Fig4_summary_updated <- ggplot() +

  # Distribution Curves (Low, Medium, High Conservation)
  stat_function(fun = dnorm, args = list(mean = H0_mean_day_of_year, sd = HC_sd), geom = "area", fill = color_curve_fill, alpha = 0.4) +
  stat_function(fun = dnorm, args = list(mean = H0_mean_day_of_year, sd = HC_sd), geom = "line", color = color_curve_line, linewidth = 0.6, alpha = 0.6) +
  stat_function(fun = dnorm, args = list(mean = H0_mean_day_of_year, sd = MC_sd), geom = "area", fill = color_curve_fill, alpha = 0.5) +
  stat_function(fun = dnorm, args = list(mean = H0_mean_day_of_year, sd = MC_sd), geom = "line", color = color_curve_line, linewidth = 0.6, alpha = 0.6) +
  stat_function(fun = dnorm, args = list(mean = H0_mean_day_of_year, sd = LC_sd), geom = "area", fill = color_curve_fill, alpha = 0.7) +
  stat_function(fun = dnorm, args = list(mean = H0_mean_day_of_year, sd = LC_sd), geom = "line", color = color_curve_line, linewidth = 0.6, alpha = 0.7) +
  
  # Add the observed data points, with color mapped to Spawn_evidence
  geom_jitter(
    data = fig4data_analysis,
    aes(x = Day_of_year, y = sample_n / scaling_factor, fill = Spawn_evidence),
    shape = 21, 
    color = "white", 
    alpha = 0.8,
    size = 4,
    stroke = 0.4
  ) +
  
  #Format
  scale_fill_manual(
    name = "Spawning Evidence",
    values = c("Yes" = color_spawn_yes, "No" = color_spawn_no)
  ) +
  
  #Scales and axes
  scale_x_continuous(limits = c(100, 400)) +
  scale_y_continuous(
    name = "Density (Normal distribution)",
    sec.axis = sec_axis(
      trans = ~ . * scaling_factor,
      name = "Number of samples (per trip)"
    )
  ) +
  
  #Titles
  labs(
    title = "Spawn observations outside of expected spawning distributions",
    subtitle = "Molokai Island only",
    x = "Day of Year"
  ) +
  
  # Apply a minimal theme and custom styling
  theme_minimal(base_family = "sans") +
  theme(
    plot.title = element_markdown(size = 14, face = "bold", margin = margin(b = 10)),
    plot.subtitle = element_text(size = 12, color = "grey50", margin = margin(b = 20)),
    plot.caption = element_text(size = 9, color = "grey60", hjust = 1),
    axis.title = element_text(size = 12, face = "bold"),
    axis.text = element_text(size = 11),
    axis.title.y.left = element_text(color = color_curve_line),
    axis.text.y.left = element_text(color = color_curve_line),
    axis.title.y.right = element_text(color = "black"), # Secondary axis title color
    axis.text.y.right = element_text(color = "black"),  # Secondary axis text color
    panel.grid.major = element_line(color = "lightgrey", size = 0.5),
    panel.grid.minor = element_blank(),
    plot.title.position = "plot",
    legend.position = "bottom" # Move legend to the bottom
  )

# Print the plot
print(Fig4_summary_updated)

# To save the plot
ggsave("10b_Fig4-plot.png", plot = Fig4_summary_updated, width = 12, height = 8, dpi = 300)


```

Manually add color and p-values to data points (done in Illustrator)

# Save all data objects into one R Workbook

```{r}
install.packages("writexl")
library(writexl)

#Named list of data frames
objects_to_save <- list(
  "10a_raw-data"= fig3data,
  "10a_all-analysis" = fig3analysis,
  "10a_p<0.001" = Fig3analysis_key,
  "10b_raw-data" = fig4data,
  "10b_all-analysis"=fig4spawn
)

#Save to excel file
write_xlsx(objects_to_save, path = "Supplementary_10ab_Data-Tables.xlsx")

```
